What We Eval in the Shadows

### 2021-04-16 paper/main.tex
Grammarly's proofreading of Section V.2 (environments)

\begin{compactitem}[---]
\item \emph{Function:} environments for the local variables of some function
  currently active on the call stack. Obtained by calling \c{parent.frame()} or
  \c{sys.frame()}.
\item \emph{Synthetic:} environments built from data structures such as lists,
  dataframes, or constructed explicitly with \c{new.env}, \c{list2env} or
  \c{as.environment}. Also includes the empty environment.
\item \emph{Global:}  environments in which scripts or interactive commands
  are evaluated.
\item \emph{Package:}  environments of a loaded library.
\end{compactitem}

\noindent
As shown in Table~\ref{tab:highlevelenvironments}, most calls evaluate in a
function environment with the \c{global} kind as a distant second. This means
that most variable lookups and most side effects either read/update existing
local variables or introduce new ones. But for which function? From the point of
view of the function that called \eval, Table~\ref{tab:funoffset} gives an
offset on the call stack. Thus 0 is the direct caller and corresponds to local variables; 1 is its parent, and so on. The data suggests that in 81\% of cases, \eval accesses its caller's environment -- this means, the variables of the function where \eval textually
occurs are read and written to. It is interesting that some 1.5\% of sites
evaluate a code in an environment that is three frames or above from \eval. This
distance implies that, in general, modular reasoning is impossible, to
understand the behavior of any piece of code entails fully understanding the
behavior of all functions that the code may call as the actions of \eval may
happen at a distance. Finally note that any particular site may have several
kinds of environments, but in \packageNbOneCategoryEnvirSitePercent of the case a
site has a single kind. It means that we can predict the \emph{kind} of the expression passed to \eval in \packageNbOneCategoryEnvirSitePercent of the sites.

@@

\noindent
Sites that evaluate in the global environment are likely split between
intentional and accidental use. Direct references to the top-level, using
\c{globalenv()} or \c{.GlobalEnv}, are rare, they occur in only
\packageNbExplicitGlobalSites sites. We suspect that the majority of uses of the
global environment come from the fact that our corpus consists of scripts that
are loaded at the top level, thus that environment may often be the caller of
\eval or close to it. The reason we make this point is that values stored in the
global environment are visible to all functions and are not reclaimed by the
garbage collector. So accidental uses may pollute that name space.

@@

\paragraph{Discussion}
The data presented in this section is what one could expect. ({\it Function})
The majority of \eval call sites access the environment of some function on the
call stack. Usually, the current function, but sometimes frames arbitrarily far
up the call stack. One thing the data does not say is how that environment was
obtained. The expected case is that the environment was the one obtained from a
promise; as promises combine code and the environment in which that code
originates. Likely less frequent are cases where \eval is provided the results
of programmatically selecting some call stack. ({\it Synthetic}) The relative
high frequency of synthetic environments corresponds to cases when one wants to
evaluate an expression in either a restricted environment, use a data
structure as an environment. ({\it Global}) The fact that the global environment
shows up so frequently is likely an artifact of the way the code is run. ({\it
  Package}) There are only 92 sites that use a package environment in \eval,
this is probably best as it is bad form to add bindings or modify functions
of a loaded package. Many more sites rather wrap a package environment in a new
environment to create singletons in a package.

>>>>

\begin{compactitem}[---]
\item \emph{Function:} environments for the local variables of some function
  currently active on the call stack. Obtained by calling \c{parent.frame()} or
  \c{sys.frame()}.
\item \emph{Synthetic:} environments built from data structures such as lists,
  dataframes, or constructed explicitly with \c{new.env}, \c{list2env}, or
  \c{as.environment}. It also includes the empty environment.
\item \emph{Global:}  environments in which scripts or interactive commands
  are evaluated.
\item \emph{Package:}  environments of a loaded library.
\end{compactitem}

\noindent
As shown in Table~\ref{tab:highlevelenvironments}, most calls evaluate in a
function environment with the \c{global} kind as a distant second. This means
that most variable lookups and most side effects either read/update existing
local variables or introduce new ones. But for which function? From the point of
view of the function that called \eval, Table~\ref{tab:funoffset} gives an
offset on the call stack. Thus 0 is the direct caller and corresponds to local variables; 1 is its parent, and so on. The data suggests that in 81\% of cases, \eval accesses its caller's environment -- this means  the variables of the function where \eval textually
occurs are read and written to. It is interesting that some 1.5\% of sites
evaluate a code in an environment that is three frames or above from \eval. This
distance implies that, in general, modular reasoning is impossible. Understanding the behavior of any piece of code entails fully understanding the
behavior of all functions that the code may call as the actions of \eval may
happen at a distance. Finally, note that any particular site may have several
kinds of environments, but in \packageNbOneCategoryEnvirSitePercent of the cases,  a site has a single kind. It means that a JIT compiler can predict the \emph{kind} of  expression passed to \eval in \packageNbOneCategoryEnvirSitePercent of the sites by just examining at the first invocation of a site.

@@

\noindent
Sites that evaluate in the global environment are likely split between
intentional and accidental use. Direct references to the top-level, using
\c{globalenv()} or \c{.GlobalEnv}, are rare; they occur in only
\packageNbExplicitGlobalSites sites. We suspect that the majority of uses of the
global environment come from the fact that our corpus consists of scripts loaded at the top level. Thus, that environment may often be the caller of
\eval or close to it. The reason we make this point is that values stored in the
global environment are visible to all functions and are not reclaimed by the
garbage collector. So accidental uses may pollute that namespace.

@@

\paragraph{Discussion}
The data presented in this section is what one could expect. ({\it Function})
The majority of \eval call sites access the environment of some function on the
call stack. Usually, the current function, but sometimes frames arbitrarily far
up the call stack. One thing the data does not say is how that environment was
obtained. The expected case is that the environment was the one obtained from a
promise, as promises combine code and the environment in which that code
originates. Likely less frequent are cases where \eval is provided the results
of programmatically selecting some call stack. ({\it Synthetic}) The relatively
high frequency of synthetic environments corresponds to cases when one wants to
evaluate an expression in either a restricted environment, use a data
structure as an environment. ({\it Global}) The fact that the global environment
shows up so frequently is likely an artifact of how the code is run. ({\it
  Package}) There are only \packageNbPackageNamespaceEnvSites sites that use a %77 sites
 package environment in \eval. This is probably best as it is an improper form of adding bindings or modifying functions of a loaded package. Many more sites instead wrap a package environment in a new
environment to create singletons in a package.
